Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept

### 2021-12-19 paper/article.tex
Paper progress

\begin{abstract}
\textit{Objective} Brain co-processors \cite{rao.coproc} are devices which use artificial
intelligence (AI) for closed-loop neurostimulation, to shape neural activity and to bridge
injured neural circuits for targeted repair and rehabilitation. The co-processor framework
offers a flexible approach to learning closed-loop stimulation policies that optimize for
(a) specific regimes of neural activity, or (b) external task performance.
For example, one may seek to learn to stimulate the motor cortex of a stroke patient,
conditioning the stimulation on upstream visual information, aiding the patient's
attempt to grasp an object. Through the use of artificial neural
networks (ANNs) and deep learning, the co-processor co-adapts with the
neural circuit, allowing it to seek optimal stimulation policies, and adapt them
as the neural circuit changes. The results presented here demonstrate a
neural co-processor for the first time, through the use of a simulation. We
explore some of the core algorithms that may allow co-processors to successfully
learn and to adapt to non-stationarity in both the brain and sensors.
\textit{Approach} We provide the first proof-of-concept of a neural co-processor that
leverages deep learning, through the use of a simulated neural circuit.
That circuit performs a reach-to-grasp task, based on visual input, and is designed to
closely resemble a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate
a variety of lesions by altering the model, and then demonstrate our co-processor's ability
to restore lost function through ``stimulation'' of that model. We further test the
ability of our co-processor to adapt its stimulation as the simulated brain undergoes changes.
\textit{Main results} Our simulated co-processor successfully co-adapts with the neural
circuit to accomplish the external reaching task. The co-processor framework
demonstrated here adapts to a variety of lesion types, and to ongoing changes in the
simulated brain.
\textit{Significance} The proof-of-concept here outlines a co-processor model, as well
as our approach to training it, leading to insights on how such a model may be
developed for \textit{in vivo} use. We believe this co-processor design will allow for
learning complex stimulation policies that help restore function to a stroke victim.
\end{abstract}

@@

While open loop stimulation techniques have yielded clinically useful results,
their results in many domains have been mixed, such as use in visual
prostheses \cite{bosking.visual}, and use in invoking somatosensory feedback
\cite{cronin.tact}. Likely this is due to the stimulation not being conditioned
on the ongoing dynamics of the circuit being stimulated. Moment-to-moment and
throughout the day, the circuit will respond differently to the same stimulus,
as a result of differing inputs and ongoing activity. Stimulation
therefore needs to be proactively adapted in response. This need is even
greater over longer time scales as the effects of plasticity and ageing change
the connectivity of the brain. Closed-loop stimulation may also provide means
to better regulate the energy use of an implanted stimulator, allowing it to
intelligently regulate when to apply stimulation, in order to presever implant
battery life. Closed-loop stimulation also offers an opportunity minimize
side-effects of stimulation, through real time regulation
of the stimulation parameters, such as in the use of deep brain stimulation (DBS) in
PD patients \cite{little.park}.

@@

For a neurologically complex task such as grasping, we cannot identify
\textit{a priori} a real time controller of the neural circuit. That is
due in large part to the variability of circuits from subject-to-subject,
as well as variations in the placement of sensors and stimulators in the
brain. The only plausible path to such a real time controller is to parameterize
it in a subject- and time-specific way. Our model seeks to accomplish that using
deep learning, together with a data-efficient approach to training.

Before attempting \textit{in vivo} experiments using such a model, we first demonstrate
here a number of crucial design elements of it, through the use of a simulated
grasping circuit, presented previously by Michaels et al. \cite{michaels.mrnn}. We
explore the properties of the artificial neural networks (ANNs) that are needed to
successfully adapt to the long-running dynamics of a stimulated neural circuit,
as well as to adapt to that circuit's ongoing connectivity changes.
We additionally present a training method our co-processor, allowing it to learn
an optimal stimulation pattern that drives improvements in external task performance,
while also adapting to the non-stationarity of the brain.

@@

\subsection{Simulation Overview}
We demonstrate our approach here using a simulated grasping circuit.
A detailed simulation such as this allows us to explore some of the critical
architectural details and training algorithms for a co-processor of this type.
By first doing such exploration in simulation, we are able to rapidly and
cheaply iterate on our design, prior to any \textit{in vivo} experiments.
For the simulation to be admissible, we need to ensure it has some
properties that allow it to strongly indicate if our design is improving in
a direction that will later allow for real deployments. Otherwise, our
design may be adapting to the pecularities of the simulation, without
becoming more useful.

An example of such a simulation approach is Dura-Bernal et al. \cite{bernal.sim}. In this work, the
authors use a simulated spiking neural network to train a stimulation agent.
Their stimulation agent sought to restore the network's
control of a simulated arm, to reach a target, after a simulated lesion was applied,
much like the work here.  As the authors point out, we have a limited ability to
probe a neural circuit \textit{in vivo} in order to perform learning. As a result,
we first need to design our approach through the use of an admissible simulation.
The authors in this case simulated lesions by effectively removing parts of their
simulated network, or by cutting connections between parts of the network.
Here, we use an established design for an artificial neural circuit which
performs a grasping task, whose architecture and training methods were designed
specifically to result in naturalistic dynamics.

@@

The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
representing the visual processing needed to reach the hand to the appropriate position for
the grasp, and to form the hand properly for grasping the particular shape of object. The
emergent dynamics of the artificial network's ``modules'', once trained, correspond roughly
to the AIP, F5, and M1 portions of the monkey subjects' brains. `Correspond' here, notably,
refers to the fact that the activity of the module receiving the visual inputs resembles
the AIP activity of the monkey subject from the same trials. Likewise, the second and
third modules resemble the natural activity from the monkey's F5 and M1 regions, respectively.
The emergent network dynamics show a number of other correspondences to the monkeys' natural
brain activity as well, as detailed in the paper.

Importantly, the simulated circuit's activity shows a relatively clear separation of the
object shapes. That is - the visual information input to the network that differentiates
one trial from another is leveraged by the circuit to properly condition
the hand shape trajectory for grasping the object of that trial's particular shape, size, and
location. The visual information forward-propagates through the network, through successive
processing steps, where it is eventually leveraged to recapitulate the grasp appropriate
for the object represented by the input visual features. It follows, then, that the classes
are separable in some way by observing the distinct ways that the mRNN is activated by the
classes' corresponding visual inputs. As we will note below: in the absense of this
visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
waiting period, but performance suffers drastically.

@@

\subsubsection{Simulated lesions cause real world failure modes}
Simulating a brain lesion in terms of this artificial network results in error
modes that resemble a some natural lesions of a primate brain. For example, if we
alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
the reaching motion generally succeeds, but the finger muscle velocities show a high degree
of error - effectively implying that the subject can somewhat reach to grasp, but cannot form a
grasp appropriate for the object. That suggests that losing a portion of the cortical
machinery needed to translate or forward-propagate object shape information
to movement-related cortex can result in a reduced ability to form the hand properly for
the grasp, though positioning of the hand may still roughly succeed. Muscle spasticity of
the hand is also a common symptom of certain strokes in primates, and indeed many who
suffer from it are still able to position their hand, even while being unable to form
it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
simulated lesion closely resembles natural stroke symptoms. However, the simulation is not
intended to constitute a physical model of a lesion, i.e. to directly explain the
connection between hand spasticity and the lesion.

@@

The co-processor's task in our simulation will be identifying the appropriate information,
read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
to effectively bridge across the lesion, forward propagating the object shape and task
structyure information, indirectly, through the use of stimulation.

In our experiments below, we demonstrate our co-processor design on three types of
simulated lesions:
\begin{itemize}
	\item \textbf{Output-based (AIP):} we force the output of some proportion of ``AIP''
	      neurons to zero, effectively removing them from the network. This results in a
	      loss of object shape information until the brain co-adapts to the lesion.
	\item \textbf{Connection-based:} we prevent the propagation of
	      information between the ``F5'' and ``M1'' modules, effectively representing a
	      severing of the connections between the two. Note that the sparse connections
	      between them run both directions, and we are lesioning both.
	\item \textbf{Output-based (M1):} we force the output of some proportion of ``M1''
	      neurons to zero, causing a more holisitic loss of movement. Here, the lesion
	      may make it impossible for the co-processor to find a solution, but some
	      recovery may be possible.
\end{itemize}

@@

To illustrate this, suppose we applied a small, one-time, instantaneous perturbation
of the hidden state of 10 randomly chosen neurons in the output (M1) module at some point
in time during a trial. If we repeat that experiment many times, we can see what the
distribution of long- running effects tends to look like.

@@

The intent with this approach is not to create a realistic biophysical model. The
mRNN model does not contain sufficient information to arrange the neurons in a
simulated space, and to simulate how that may change the effects of stimulation.
Instead, our stimulation model is merely attempting to make the co-processor's task
more realistic, in the sense that it cannot precisely set the hidden state of every
neuron, instantaneously and in each time step. Any stimulation it applies will diffuse
across ``space'' and time.

@@

Likewise, the stimulation function maps the relatively low dimensional stimulation
parameter vector $\theta$ onto the M1 cells which it stimulates. In our experiments,
our stimulation parameters $\theta$ had 16 dimensions, which we found was
sufficient to allow significant improvement in task performance. The stimulation
function uses a simple Gaussian smoothing to map those parameters onto the simulated
M1 neurons. Each parameter represents, in a sense, an electrode located along a single
spatial dimension, whose stimulation affects the neurons in its vicinity more than
it affects others, according to each neuron's Mahalanobis distance from it. The neurons
are aligned along that dimension arbitrarily. We fix the width of the Gaussians
arbitrarily to $\sigma=1.75$. We are not aware of a principled way of picking this
value, and clearly if we set it to extreme values, it makes our simulation less useful.
For example, if it is set to an extreme high value, the stimulation is effectively
lower dimensional, resulting in a model corresponding to a stimulation technology that
is inappropriately low resolution for our task, for example epidural stimulation.
However, we cannot conclude much from these experiments, because this simulation does
not constitute a biophysical model, where we can identify some realistic value. The
purpose of this smoothing technique is only to make the co-processor's job harder in
the simple sense that it cannot target neurons individually.
As a result, we set it to a moderate, balanced value, and don't vary it.

@@

See Fig. \ref{fig:stim_single} for a visual depiction of an example where
$\theta$ is non-zero at $t=0$, and zero for all other times, in order to see how
stimulation is applied across space and time. In Fig. \ref{fig:stim_and_obs},
we see stimulation applied during the simulation, in a real trial. In this case,
the trial is one where $50\%$ of M1 neurons are lesioned to have zero output.
Observations of the network's hidden state (explained in the next section) are taken
from the first two modules, and stimulation is applied to the last module. This fact
is intended to capture the difficulties with observing the same region that you are
stimulating, due to stimulation artifacts. The trial we show depicts stimulation
supplied by a highly trained co-processor, where task performance has been
largely restored. We explain this experiment in more detail below.

@@

We achieve this through a simple error backpropagation, using PyTorch's
implementation of the $Adam$ optimizer. With each trial, task loss is
calculated and backpropagated into the mRNN, allowing it to adapt to
the fact it receives stimulation. The learning rate for the optimizer
is set arbitrarily to a relatively low rate of $1e-7$. The learning
rate was not chosen in a principled way, since we are not aware of a
principled way of choosing it. However, we reason that it should be
lower than the co-processor by orders of magnitude due
to the rapidity with which an ANN can be retrained.

\subsubsection{Simulating recovery prior to co-processor use}
We sought to additonally re-train the mRNN after applying the lesion in
order to simulate stroke recovery. For our simulated lesions which zero
the outputs of neurons, the mRNN cannot supply an appropriate pre-recovery model.
We know that strokes can cause damage from which the subject, in many cases,
cannot recover. The mRNN model has sufficient redundancy built into it that
lesioning the outputs of large numbers of neurons leaves enough remaining
degrees of freedom that a nearly full recovery can occur, unless so many
neurons are lesioned that no stimulation could be effective.

@@

\subsubsection{EN training}
For the EN to be useful, it must accurately predict the effect of stimulation
produced by the CPN. In addition to that, backpropagating through it must yield
gradients which train the CPN to output more useful stimulation. We discovered that
this latter property does not occur simply by virtue of the former.
We have found that an EN can be trained to high levels of predictive power, even
on random stimulation, and to orders of magnitude lower loss than the task loss,
while at the same time backpropagation through it yields gradients which are not
useful for training the CPN. As a result, if the EN is not trained in specific ways,
the CPN exhibits unstable training. We have thus far been unable to discover a way
to measure and optimize for this second property directly, though pursuing it may be a
fruitful subject of future applied ML research. Instead of attempting to train an EN
directly to exhibit this property, we discovered a training regime that indirectly,
but reliably, makes it occur.

@@

The choice of learning rate schedule for training the CPN is somewhat important.
The CPN appears to train in two phases. In the first phase, it is largely learning the structure
of the reach-to-grasp task, e.g. that the muscles need to stay still until the reach begins. During
that phase, the learning rate can be quite high. Later, the CPN begins to learn the mapping between
the object shape information observed from the simulated brain, and how that should map onto
stimulation. That phase of the training takes much longer, and requires a learning rate 2-3
orders of magnitude lower than in the first phase. We will cover this in some more detail in
Section \ref{sec:results}.

@@

In the co-adaptive case, the brain and co-processor find a solution that largely recovers
performance. This demonstrates that the co-processor can adapt to the non-stationarity
of the mapping between the observed brain activity and the optimal stimulation. Note,
though, that the simulation is not designed to indicate if the co-processor speeds up
recovery, or results in greater results than natural recovery, since the simulation
is not designed to answer that question.

@@

\subsection{Experiment 2: M1 75\% lesion}
In this experiment, 75\% of the output module of the simulated network has been
lesioned. In the coadaptive case, the co-processor needs to adapt to
non-stationarity in the mapping between stimulation and task performance.
In the non-coadaptive case, training quickly plateaus, and does not improve
thereafter. That suggests we have lost some of our degrees of freedom to control
the activations in the output layer, resulting in a reduced ability to recreate
the target grasps.

In the co-adapative case, performance likewise hits an inflexion point, but
continues to improve slowly.  TODO: finish once we have the faster co-adaptation
rate tested.

\subsection{Experiment 3: F5 and M1 connectivity lesion}
As in the other cases, the co-processor's loss in this experiment improves
quickly, initially, but taskes much longer to refine. During that refinement,
object differentiation occurs. Note that in this experiment, no information
can propagate forward to the M1 module, since we have completely severed the
input connections to that module. It is the sole responsibility of the
co-processor to perform that forward-propagation.

In this case, co-adaptation only affects the M1 module, since we have nothing
to use as a learning signal: there is no backpropagation to the AIP and
F5 modules. As a result, this experiment primarily demonstrates that the
co-processor's learning algorithm is capable of adapting to non-stationarity
in the mapping between stimulation parameters and task performance. Contrast
that with the results of Experiment 1, where the non-stationarity affected
the mapping between the visual inputs and the brain observations.

@@

The co-processor in this scenario successfully improved task performance.
Note that the object class separation metric $S$ returned to
a healthy value towards the end of the training, indicating that the
learned solution did not simply refine the stereotyped grasp.

@@

Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
50-90\% across our various experiments. To work this well, the co-processor needed to adapt to
the long-running dynamics of the stimulated network, as well as the long-running effects of
stimulation. In some experiments we required it to additionally adapt to non-stationarity in
the neural circuit, which was actively changing at the same time the co-processor was learning.
It achieved that as well, co-adapting with the simulated brain to improve external task
performance. Finally, it adapted to a simulated brain which had already undergone some amount of
lesion recovery. In that case, the co-processor successfully identified the information upstream
from the lesion which was necessary to stimulate the motor cortex downstream.

@@

Abstractly, the co-processor lends itself to any closed-loop neural stimulation problem
where a relevant stimulation model can be identified. There are many lower dimensional problems
today which we address with far simpler stimulation regimes than would be necessary for
e.g. remediating the symptoms of a stroke. In some applications, such as PD symptom relief,
stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
parameters, where the key challenge is learning when to apply the stimulation, and
its shape and power. Our simulation demonstrates the ability of our training method to adapt to
higher dimensional problems than these, where additionally the long-running effects of
stimulation on the intended figure-of-merit must be modeled. These properties together
suggest it may be useful in lower dimensional problems where a stimulation model can be
estimated, and stimulation parameters can be tied to the target objective.

@@

\section{Conclusion}
We have demonstrated a general framework for training a closed-loop neural stimulator,
called a ``neural co-processor''. We outlined one specific co-processor design, and
outlined an approach to training it using supervised learning. In simulation it
successfully learned to alleviate some of the effects of a brain lesion. To do so, it
needed to co-adapt to the simulated brain, which exhibited non-stationarity. Additionally,
it needed to learn long-running effects of the stimulation it learned to apply. Because of
these properites, the co-processor approach likely generalizes well to a wide range of
clinical applications where closed-loop neural stimulation is appropriate.

>>>>

\begin{abstract}
\textit{Objective} Brain co-processors \cite{rao.coproc} are devices which use artificial
intelligence (AI) for closed-loop neurostimulation, to shape neural activity and to bridge
injured neural circuits for targeted repair and rehabilitation. The co-processor framework
offers a flexible approach to learning closed-loop stimulation policies that optimize for
(a) specific regimes of neural activity, or (b) external task performance.
For example, one may seek to learn to stimulate the motor cortex of a stroke patient,
conditioning the stimulation on upstream visual information, aiding the patient's
attempt to grasp an object. Through the use of artificial neural
networks (ANNs) and deep learning, the co-processor co-adapts with the
neural circuit, allowing it to seek optimal stimulation policies, and adapt them
as the neural circuit changes. The results presented here demonstrate a
neural co-processor for the first time, through the use of a simulation. We
explore some of the core algorithms that may allow co-processors to successfully
learn, and to adapt to non-stationarity in both the brain and sensors.
\textit{Approach} We provide the first proof-of-concept of a neural co-processor that
leverages deep learning, through the use of a simulated neural circuit.
That circuit performs a reach-to-grasp task, based on visual input, and is designed to
closely resemble a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate
a variety of lesions by altering the model, and then demonstrate our co-processor's ability
to restore lost function through ``stimulation'' of that model. We further test the
ability of our co-processor to adapt its stimulation as the simulated brain undergoes changes.
\textit{Main results} Our simulated co-processor successfully co-adapts with the neural
circuit to accomplish the external reaching task. The co-processor framework
demonstrated here adapts to a variety of lesion types, and to ongoing changes in the
simulated brain.
\textit{Significance} The proof-of-concept here outlines a co-processor model, as well
as our approach to training it, leading to insights on how such a model may be
developed for \textit{in vivo} use. We believe this co-processor design has potential to
learn complex stimulation policies, leading to application in treating a variety of
diseases.
\end{abstract}

@@

While open loop stimulation techniques have yielded clinically useful results,
their results in many domains have been mixed, such as use in visual
prostheses \cite{bosking.visual}, and use in invoking somatosensory feedback
\cite{cronin.tact}. Likely this is due to the stimulation not being conditioned
on the ongoing dynamics of the circuit being stimulated. Moment-to-moment and
throughout the day, the circuit will respond differently to the same stimulus,
as a result of differing inputs and ongoing activity. Stimulation
therefore needs to be proactively adapted in response. This need is even
greater over longer time scales as the effects of plasticity and ageing change
the connectivity of the brain. Closed-loop stimulation may also provide means
to better regulate the energy use of an implanted stimulator, allowing it to
intelligently regulate when to apply stimulation, in order to preserve implant
battery life. Closed-loop stimulation also offers an opportunity minimize
side-effects of stimulation, through real time regulation
of the stimulation parameters, such as in the use of deep brain stimulation (DBS) in
PD patients \cite{little.park}.

@@

For a neurologically complex task such as grasping, we cannot identify
\textit{a priori} a real time controller of the neural circuit. That is
due in large part to the variability of circuits from subject-to-subject,
as well as variations in the placement of sensors and stimulators in the
brain. The only plausible path to such a real time controller is to parameterize
it in a subject- and time-specific way. Our model seeks to accomplish that using
deep learning.

Before attempting \textit{in vivo} experiments using such a model, we first demonstrate
here a number of crucial design elements of it, through the use of a simulated
grasping circuit, presented previously by Michaels et al. \cite{michaels.mrnn}. We
explore the properties of the co-processor design that are needed to
successfully adapt to the long-running dynamics of a stimulated neural circuit,
as well as to adapt to that circuit's ongoing connectivity changes.
We additionally present a training method for our co-processor, allowing it to learn
an optimal stimulation pattern that drives improvements in external task performance,
while also adapting to the non-stationarity of the brain.

@@

\subsection{Simulation Overview}
We demonstrate our approach here using a simulated grasping circuit.
A detailed simulation such as this allows us to explore some of the critical
architectural details and training algorithms that will allow a co-processor
to succeed.  By first doing such exploration in simulation, we are able to
rapidly and cheaply iterate on our design, prior to any \textit{in vivo}
experiments. For the simulation to be admissible, we need to ensure it has some
properties that allow it to strongly indicate if our design is improving in
a direction that will later allow for real deployments. Otherwise, our
design may be adapting to the pecularities of the simulation, without
becoming more useful.

An example of such a simulation approach is Dura-Bernal et al. \cite{bernal.sim}. In this work, the
authors use a simulated spiking neural network to train a stimulation agent.
Their stimulation agent sought to restore the network's
control of a simulated arm, to reach a target, after a simulated lesion was applied,
much like the work here.  As the authors point out, we have a limited ability to
probe a neural circuit \textit{in vivo} in order to perform learning. As a result,
we first need to design our approach through the use of an admissible simulation.
The authors in this case simulated lesions by effectively removing parts of their
simulated network, or by cutting connections between parts of the network.
Here, we use an established design for an artificial neural circuit which
performs a grasping task. Its architecture and training methods were designed
specifically to result in naturalistic dynamics.

@@

The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
representing the visual processing needed to reach the hand to the appropriate position for
the grasp, and to form the hand properly for grasping the particular shape of object. The
emergent dynamics of the artificial network's ``modules'', once trained, correspond roughly
to the AIP, F5, and M1 portions of the monkey subjects' brains. `Correspond' here, notably,
refers to the fact that the activity of the module receiving the visual inputs resembles
the AIP activity of the monkey subject from the same trials. Likewise, the second and
third modules resemble the natural activity from the monkey's F5 and M1 regions, respectively.
The emergent network dynamics show a number of other correspondences to the monkeys' natural
brain activity as well, as detailed in the paper. For convenience, we will refer to these
modules throughout this paper using the brain region they correspond to.

Importantly, the simulated circuit's activity shows a relatively clear separation of the
object shapes. That is - the visual information input to the network that differentiates
one trial from another is leveraged by the circuit to successfully condition
the hand shape trajectory for grasping the object of that trial's particular shape, size, and
location. The visual information forward-propagates through the network, through successive
processing steps, where it is eventually leveraged to recapitulate the grasp appropriate
for the object represented by the input visual features. It follows, then, that the classes
are separable in some way by observing the distinct ways that the mRNN is activated by the
classes' corresponding visual inputs. As we will note below: in the absense of this
visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
waiting period, but performance suffers drastically.

@@

\subsubsection{Simulated lesions cause real world failure modes}
Simulating a brain lesion in terms of this artificial network results in error
modes that resemble a some natural lesions of a primate brain. For example, if we
alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
the reaching motion generally succeeds, but the finger muscle velocities show a high degree
of error - effectively implying that the subject can somewhat reach to grasp, but cannot form a
grasp appropriate for the object. That suggests that losing a portion of the cortical
machinery needed to translate or forward-propagate object shape information
to movement-related cortex can result in a reduced ability to pose the hand, though
positioning of the hand may still roughly succeed. Muscle spasticity of
the hand is also a common symptom of certain strokes in primates, and indeed many who
suffer from tem are still able to position their hand, even while being unable to form
it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
simulated lesion closely resembles natural stroke symptoms. However, the simulation is
not intended to constitute a physical model of a lesion, i.e. to directly explain the
connection between hand spasticity and the lesion.

@@

The co-processor's task in our simulation will be identifying the appropriate information,
read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
to effectively bridge across the lesion, forward propagating the object shape and task
structure information, indirectly, through the use of stimulation.

In our experiments below, we demonstrate our co-processor design on three types of
simulated lesions:
\begin{itemize}
	\item \textbf{Output-based (AIP):} we force the output of some proportion of ``AIP''
	      neurons to zero, effectively removing them from the network. This results in a
	      loss of object shape information.
	\item \textbf{Connection-based:} we prevent the propagation of
	      information between the ``F5'' and ``M1'' modules, effectively representing a
	      severing of the connections between the two. Note that the sparse connections
	      between them run both directions, and we are lesioning both.
	\item \textbf{Output-based (M1):} we force the output of some proportion of ``M1''
	      neurons to zero, causing a more holisitic loss of movement. Here, the lesion
	      may make it impossible for the co-processor to find a solution, but some
	      recovery may be possible.
\end{itemize}

@@

To illustrate this, suppose we applied a small, one-time, instantaneous perturbation
of the hidden state of 10 randomly chosen neurons in the output (M1) module at some point
in time during a trial. If we repeat that experiment many times, we can see what the
distribution of long-running effects tends to look like on the output muscle velocities.

@@

The intent with this approach is not to create a realistic biophysical model. The
mRNN model does not contain sufficient information to arrange the neurons in a
simulated space, and to simulate how that may change the effects of stimulation.
Instead, our stimulation model is merely attempting to make the co-processor's task
more realistically difficult, in the sense that it cannot precisely set the hidden
state of every neuron, instantaneously and in each time step. Any stimulation it
applies will diffuse across ``space'' and time.

@@

Likewise, the stimulation function maps the relatively low dimensional stimulation
parameter vector $\theta$ onto the M1 cells which it stimulates. In our experiments,
our stimulation parameters $\theta$ had 16 dimensions, which we found was
sufficient to allow significant improvement in task performance. The stimulation
function uses a simple Gaussian smoothing to map those parameters onto the simulated
M1 neurons. Each parameter represents, in a sense, an electrode located along a single
spatial dimension, whose stimulation affects the neurons in its vicinity more than
it affects others, according to each neuron's Mahalanobis distance from it. The neurons
are aligned along that dimension arbitrarily. We fix the width of the Gaussians
arbitrarily to $\sigma=1.75$. We are not aware of a principled way of picking this
value, and clearly if we set it to extreme values, it makes our simulation less useful.
For example, if it is set to an extreme high value, the stimulation is effectively
lower dimensional, resulting in a model corresponding to a stimulation technology that
is inappropriately low resolution for our task, for example epidural stimulation.
We could not conclude much from such an experiment, because this simulation does
not constitute a biophysical model, where we can identify some realistic value. The
purpose of this smoothing technique is only to make the co-processor's job harder in
the simple sense that it cannot target neurons individually. As a result, we set it
to a moderate, balanced value, and don't vary it.

@@

See Fig. \ref{fig:stim_single} for a visual depiction of an example where
$\theta$ is non-zero at $t=0$, and zero for all other times, in order to see how
stimulation is applied across space and time. In Fig. \ref{fig:stim_and_obs},
we see stimulation applied during the simulation, in a real trial. In this case,
the trial is one where $50\%$ of M1 neurons are lesioned to have zero output.
Observations of the network's hidden state (explained in the next section) are taken
from the first two modules, and stimulation is applied to the last module. This fact
is intended to capture the difficulties with observing the same region that you are
stimulating, due to stimulation artifacts. The trial we show depicts stimulation
supplied by a highly trained co-processor, where task performance has been
significantly restored. We explain this experiment in more detail below.

@@

We achieve this through a simple error backpropagation, using PyTorch's
implementation of the $Adam$ optimizer. With each trial, task loss is
calculated and backpropagated into the mRNN, allowing it to adapt to
the fact it receives stimulation. The learning rate for the optimizer
is set arbitrarily to a relatively low rate of $1e-7$. The learning
rate was not chosen in a principled way, since we are not aware of a
principled way of choosing it. However, we reason that it should be
lower than the co-processor by orders of magnitude due
to the rapidity with which an ANN can be retrained in practice.

\subsubsection{Simulating recovery prior to co-processor use}
We sought to additonally re-train the mRNN after applying the lesion in
order to simulate stroke recovery. For our simulated lesions which zero
the outputs of neurons, the mRNN cannot supply an appropriate pre-recovery model.
We know that strokes can cause damage from which the subject, in many cases,
cannot recover. The mRNN model has sufficient redundancy built into it that
lesioning the outputs of large numbers of neurons often leaves enough remaining
degrees of freedom that a nearly full recovery can occur, unless so many
neurons are lesioned that no stimulation could be effective.

@@

\subsubsection{EN training}
For the EN to be useful, it must accurately predict the effect of stimulation
produced by the CPN. In addition to that, backpropagating through it must yield
gradients which train the CPN to output more useful stimulation. We discovered that
this latter property does not occur simply by virtue of the former.
An EN of this design can be trained to high levels of predictive power, even
on random stimulation, and to orders of magnitude lower loss than the task loss,
while at the same time backpropagation through it yields gradients which are not
useful for training the CPN. As a result, if the EN is not trained in specific ways,
the CPN exhibits unstable training. We have thus far been unable to discover a way
to measure and optimize for this second property directly, though pursuing it may be a
fruitful subject of future applied ML research. Instead of attempting to train an EN
directly to exhibit this property, we discovered a training regime that indirectly,
but reliably, makes it occur.

@@

The choice of learning rate schedule for training the CPN is somewhat important.
The CPN appears to train in two phases. In the first phase, it is largely learning the structure
of the reach-to-grasp task, e.g. that the muscles need to stay still until the reach begins. During
that phase, the learning rate can be quite high. Later, the CPN begins to learn the mapping between
the object shape information observed from the simulated brain, and how that should map onto
stimulation. That phase of the training takes much longer, and requires a learning rate 2-3
orders of magnitude lower than in the first phase. We will cover this in some more detail in
the Results section (Section \ref{sec:results}).

@@

In the co-adaptive case, the brain and co-processor find a solution that largely recovers
performance. This demonstrates that the co-processor can adapt to the non-stationarity
of the mapping between the observed brain activity and the optimal stimulation. Note,
though, that the simulation is not designed to indicate if the co-processor speeds up
recovery, or results in greater results than natural recovery>

@@

\subsection{Experiment 2: M1 75\% lesion}
In this experiment, 75\% of the output module of the simulated network has been
lesioned. In the co-adaptive case, the co-processor needs to adapt to
non-stationarity in the mapping between stimulation and task performance.
In the non-coadaptive case, training quickly plateaus, and does not improve
thereafter. That suggests we have lost some of our degrees of freedom to control
the activations in the output layer, resulting in a reduced ability to recreate
the target grasps.

In the co-adapative case, performance likewise hits an inflexion point, but
continues to improve slowly. The ability to improve further is tied to the
simulated recovery restoring the degrees of freedom needed for success. The
co-processor in this later phase of training is tasked with adapting to the
non-stationarity that results from the recovery process, which it does
successfully.

\subsection{Experiment 3: F5 and M1 connectivity lesion}
As in the other cases, the co-processor's loss in this experiment improves
quickly, initially, but taskes much longer to refine. During that refinement,
object differentiation occurs. Note that in this experiment, no information
can propagate forward to the M1 module, or backwards from it, since we have
completely severed the inter-module connections involving that module. It is
the sole responsibility of the co-processor to perform that
forward-propagation.

In this case, co-adaptation only affects the M1 module, since we have nothing
to use as a learning signal: there is no backpropagation to the AIP and
F5 modules. As a result, this experiment demonstrates that the
co-processor's learning algorithm is capable of adapting to non-stationarity
in the mapping between stimulation parameters and task performance. Contrast
that with the results of Experiment 1, where the non-stationarity affected
the mapping between the visual inputs and the brain observations.

@@

The co-processor in this scenario successfully improved task performance
beyond the recovery point. Note that the object class separation metric
$S$ returned to a healthy value towards the end of the training,
indicating that the learned solution did not simply refine the
stereotyped grasp.

@@

Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
75-90\% across our various experiments. To work this well, the co-processor needed to adapt to
the long-running dynamics of the simulated network, as well as the long-running effects of
stimulation. In some experiments we required it to additionally adapt to non-stationarity in
the neural circuit, which was actively changing at the same time the co-processor was learning.
It achieved that as well, co-adapting with the simulated brain to improve external task
performance. Finally, it adapted to a simulated brain which had already undergone some amount of
lesion recovery. In that case, the co-processor successfully identified the information upstream
from the lesion which was necessary to stimulate the motor cortex downstream.

@@

Abstractly, the co-processor lends itself to any closed-loop neural stimulation problem
where a relevant stimulation model can be identified. There are many lower dimensional problems
today which we address with far simpler stimulation regimes than would be necessary for
e.g. fine-grained finger control. In some applications, such as PD symptom relief,
stimulation is often parameterized by a single off-vs-on parameter, or a small handful of
parameters, where the key challenge is learning when to apply the stimulation, and
its shape and power. Our simulation demonstrates the ability of our training method to
adapt to higher dimensional problems than these, where additionally the long-running
effects of stimulation on the intended figure-of-merit must be modeled. These properties
together suggest it may be useful in lower dimensional problems where a stimulation model
can be estimated, and stimulation parameters can be tied to the target objective.

@@

\section{Conclusion}
We have demonstrated a general framework for training a closed-loop neural stimulator,
called a ``neural co-processor''. We outlined one specific co-processor design, and
outlined an approach to training it using supervised learning. In simulation it
successfully learned to alleviate some of the effects of a brain lesion. To do so, it
needed to co-adapt with the simulated brain, which exhibited non-stationarity. Additionally,
it needed to learn long-running effects of the stimulation it learned to apply. Because of
these properites, the co-processor approach likely generalizes well to a wide range of
clinical applications where closed-loop neural stimulation is appropriate.
